BIP-Predictor

### 2025-05-10 Sweave/BIP_predictor.Rnw
up

# Plot BIP and diff-log BIP
file = "./Figures/data.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
par(mfrow=c(1,1))

# Plot the data
# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
mplot<-x_mat
colnames(mplot)[1]<-"GDP"
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Data: standardized diff-log, Pandemic outliers trimmed"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
dev.off()

@@

par(mfrow=c(1,2))
# Plot the data
# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
mplot<-x_mat[which(rownames(x_mat)>=2007&rownames(x_mat)<=2011),]
colnames(mplot)[1]<-"GDP"
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Financial Crisis"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
mplot<-x_mat[which(rownames(x_mat)>=2019&rownames(x_mat)<=2022),]
colnames(mplot)[1]<-"GDP"
colo<-c("black",rainbow(ncol(data)-1))
main_title<-"Pandemic"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
mtext(colnames(mplot)[1],col=colo[1],line=-1)
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()

@@

  }
  mplot_with_pandemic<-mplot
  ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
  for (k in 1:length(select_vec_i))
  { 
    if (k==1)
    {
      mtext("GDP",line=-k,col=colo_i[k])
    } else
    {
      mtext(select_vec_i[k],line=-k,col=colo_i[k])
    }
  }
}

@@

  }
    mplot_without_pandemic<-mplot

@@

file = "./Figures/ma_inv_BIP.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 9, height = 6)
par(mfrow=c(1,2))
ts.plot(mplot_with_pandemic,col=colo_i,main=paste("MA inversion GDP: with pandemic",sep=""),xlab="")
  for (k in 1:length(select_vec_i))
  { 
    if (k==1)
    {
      mtext("GDP",line=-k,col=colo_i[k])
    } else
    {
      mtext(select_vec_i[k],line=-k,col=colo_i[k])
    }
  }
ts.plot(mplot_without_pandemic,col=colo_i,main="Without pandemic",xlab="")
#  for (k in 1:length(select_vec_i))
#    mtext(select_vec_i[k],line=-k,col=colo_i[k])

@@

# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
for (h in 0:6)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  sum_obj<-sum_obj_unfiltered<-summary(lm_obj)
  MSE_unfiltered_vec<-c(MSE_unfiltered_vec,sum_obj$sigma^2)

@@

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  rownames(mplot)<-rownames(x_mat_wc)
  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("GDP shifted  by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    if (h==0)
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
  axis(2)
  box()
}
dev.off()

@@

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  rownames(mplot)<-rownames(x_mat_wc)
  mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    if (h==0)
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}
dev.off()

@@

par(mfrow=c(2,2))
f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected filtered indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj_hp<-summary(lm_obj)
  sum_obj<-summary(lm_obj)
  MSE_hp_vec<-c(MSE_hp_vec,sum_obj$sigma^2)
  p_val_hp<-c(p_val_hp,pf(sum_obj$fstatistic["value"], sum_obj$df[1],sum_obj$df[2], lower.tail = FALSE))
  f_vec_hp<-c(f_vec_hp,sum_obj$fstatistic["value"])
  t_mat_hp<-rbind(t_mat_hp,sum_obj$coef[-1,4])
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,h+1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,h+1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

@@

  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","blue")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    if (h==0)
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}
colnames(t_mat_hp)<-select_vec_multi
rownames(t_mat_hp)<-paste("h=",0:3)
dev.off()

if (F)
{
# rRMSEs are computed for h=0:6 (just increase h in the above lopp to recompute)
# Compare MSEs and p-values
  rrmse_hp_direct<-sqrt(MSE_hp_vec/MSE_unfiltered_vec)
  names(rrmse_hp_direct)<-names(p_val_hp)<-names(p_val_unfiltered)<-paste("h=",0:6,sep="")

  save(rrmse_hp_direct,file=paste(path.result,"/rrmse_hp_direct",sep=""))
  save(p_val_unfiltered,file=paste(path.result,"/p_val_unfiltered",sep=""))
  save(p_val_hp,file=paste(path.result,"/p_val_hp",sep=""))
} else
{
  load(file=paste(path.result,"/rrmse_hp_direct",sep=""))
  load(file=paste(path.result,"/p_val_unfiltered",sep=""))
  load(file=paste(path.result,"/p_val_hp",sep=""))

}
rrmse_hp_direct

# Financial crisis
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)

@@

for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,h+1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj$fstatistic  
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,h+1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,h+1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

@@

  colo<-c("black","blue","red")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
    if (h==0)
    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  box()
}

dev.off()
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
cat("\\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.", sep = "")
cat("\\label{direct_hp_forecasts}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
cat("\\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).", sep = "")
cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).

@@

mat1<-matrix(round(rrmse_hp_direct[4:7],5),nrow=1)
colnames(mat1)<-names(rrmse_hp_direct)[4:7]
rownames(mat1)<-"rRMSE"
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("Relative root mean-square error of HP-C against classic direct forecast: numbers smaller one signify an outperformance by the direct HP-C forecast.  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@

mat1<-matrix(round(p_val_unfiltered,5),nrow=1)
colnames(mat1)<-names(p_val_unfiltered)
rownames(mat1)<-"P-values"
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("Statistical significance of direct forecasts.  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@

@@

par(mfrow=c(1,2))
mplot<-gammak
colnames(mplot)[1]<-"GDP"
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-MSE",xlab="")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)
mplot<-bk
colnames(mplot)[1]<-"GDP"

colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA",xlab="")
#for (i in 1:ncol(mplot))
#  mtext(colnames(mplot)[i],col=colo[i],line=-i)
#mplot<-bk_HP_HP_C
#colo<-rainbow(ncol(mplot))
#ts.plot(mplot,col=colo,main="M-SSA larger HT")
#for (i in 1:ncol(mplot))
#  mtext(colnames(mplot)[i],col=colo[i],line=-i)

@@

par(mfrow=c(1,2))
# h=0, see above
mplot<-bk
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA: h=0",xlab="")
for (i in 1:ncol(mplot))
  mtext(colnames(mplot)[i],col=colo[i],line=-i)

# h=4
bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
colnames(bk)<-colnames(gammak)<-select_vec_multi
mplot<-bk
colo<-rainbow(ncol(mplot))
ts.plot(mplot,col=colo,main="M-SSA: h=4",xlab="")
#for (i in 1:ncol(mplot))
#  mtext(colnames(mplot)[i],col=colo[i],line=-i)

>>>>

# Plot BIP and diff-log BIP
file = "./Figures/data.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 6, height = 6)
par(mfrow=c(1,1))
colo<-c("black", "red", "green", "blue", "orange", "dimgray")
mplot<-x_mat
colnames(mplot)[1]<-"GDP"
main_title<-"Data: standardized diff-log, Pandemic outliers trimmed"
plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")

@@

par(mfrow=c(1,2))
# Plot the data
# The real-time BIP (red) is lagging the target (black) by lag_vec[1] quarters (publication lag)
mplot<-x_mat[which(rownames(x_mat)>=2007&rownames(x_mat)<=2011),]
colnames(mplot)[1]<-"GDP"
plot(mplot[,1],main="Financial crisis",axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
#legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
mplot<-x_mat[which(rownames(x_mat)>=2019&rownames(x_mat)<=2022),]
colnames(mplot)[1]<-"GDP"
plot(mplot[,1],main="Pandemic",axes=F,type="l",xlab="",ylab="",col=colo[1],lwd=1,ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,4*1:(nrow(mplot)/4)),labels=rownames(mplot)[c(1,4*1:(nrow(mplot)/4))])
axis(2)
box()
#legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")

@@

  }
  mplot_with_pandemic<-mplot
  colnames(mplot_with_pandemic)<-c("GDP",select_vec_i[2:length(select_vec_i)])
  ts.plot(mplot,col=colo_i,main=paste("MA inversion ",colnames(data_fit)[i],sep=""),xlab="lag")
  for (k in 1:length(select_vec_i))
  { 
    if (k==1)
    {
      mtext("GDP",line=-k,col=colo_i[k])
    } else
    {
      mtext(select_vec_i[k],line=-k,col=colo_i[k])
    }
  }
}

@@

  }
    mplot_without_pandemic<-mplot
  colnames(mplot_without_pandemic)<-c("GDP",select_vec_i[2:length(select_vec_i)])

@@

file = "./Figures/ma_inv_BIP.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 9, height = 6)
colo<-c("black", "red", "green", "blue", "orange", "dimgray")
par(mfrow=c(1,2))
plot(mplot_with_pandemic[,1],main=paste("MA inversion GDP: with pandemic",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot_with_pandemic)),max(na.exclude(mplot_with_pandemic))))

for (i in 1:ncol(mplot_with_pandemic))
{
  lines(mplot_with_pandemic[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
axis(2)
legend("topright", legend = colnames(mplot_with_pandemic), col = colo, lty = 1, lwd = 1, bty = "n")
box()

plot(mplot_without_pandemic[,1],main=paste("Without pandemic",sep=""),axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot_without_pandemic)),max(na.exclude(mplot_without_pandemic))))

for (i in 1:ncol(mplot_without_pandemic))
{
  lines(mplot_without_pandemic[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=1:nrow(mplot),labels=1:nrow(mplot)-1)
axis(2)
box()

@@

# Note: too complex designs (too many indicators) lead to overfitting and thus worse out-of-sample performances
# To illustrate the direct predictor consider the following example of a h-step ahead direct forecast:
f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),select_direct_indicator])
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  summary(lm_obj)
  sum_obj<-sum_obj_unfiltered<-summary(lm_obj)
  MSE_unfiltered_vec<-c(MSE_unfiltered_vec,sum_obj$sigma^2)

@@

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  rownames(mplot)<-rownames(x_mat_wc)
  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("GDP shifted  by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
  }
  abline(h=0)
  axis(1,at=c(1,12*1:(nrow(mplot)/12)),labels=rownames(mplot)[c(1,12*1:(nrow(mplot)/12))])
  axis(2)
      if (h==0)
    legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")

  box()
}
dev.off()

@@

  # We can now plot target and direct forecast: for h>2 the predictor comes close to a flat line centered at zero
  mplot<-cbind(forward_shifted_BIP,direct_forecast)
  rownames(mplot)<-rownames(x_mat_wc)
  mplot<-mplot[which(rownames(mplot)>2008&rownames(mplot)<2011),]
  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","red")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
#    if (h==0)
#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
    if (h==0)
    legend("bottomright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")

  box()
}
dev.off()

@@

par(mfrow=c(2,2))
f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected filtered indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj_hp<-summary(lm_obj)
  sum_obj<-summary(lm_obj)
  MSE_hp_vec<-c(MSE_hp_vec,sum_obj$sigma^2)
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

@@

  colnames(mplot)<-c("GDP","Direct forecast")
  colo<-c("black","blue")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
#    if (h==0)
#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  if (h==0)
    legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")

  box()
}
#colnames(t_mat_hp)<-select_vec_multi
#rownames(t_mat_hp)<-paste("h=",0:3)
dev.off()

# Same but financial crisis only
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
pdf(file = paste(path.out,file,sep=""), paper = "special", width = 7.5, height = 6)

@@

for (h in 0:3)
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
  # You will probably not find statistically significant regressors for h>2: BIP is a noisy series
  sum_obj<-summary(lm_obj)
  sum_obj$fstatistic  
  # Technical note: 
  # -Residuals are subject to heteroscedasticity (crises) and autocorrelation
  # -Therefore classic OLS tests for statistical significance are biased
  # -We shall rely on HAC-adjusted p-values further down (R-package sandwich)

  # Compute the predictor: one can rely on the generic R-function predict or compute the predictor manually
  direct_hp_forecast<-lm_obj$coef[1]+t(hp_c_array_wc[,,1])%*%lm_obj$coef[2:(ncol(t(hp_c_array_wc[,,1]))+1)]
  # Note that this is a full-sample predictor (no out-of-sample span)

@@

  colo<-c("black","blue","red")
  main_title<-paste("GDP shifted forward by ",h," quarters",sep="")
  plot(mplot[,1],main=main_title,axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
  for (i in 1:ncol(mplot))
  {
    lines(mplot[,i],col=colo[i])
#    if (h==0)
#    mtext(colnames(mplot)[i],col=colo[i],line=-i)
  }
  abline(h=0)
  axis(1,at=1:nrow(mplot),labels=rownames(mplot))
  axis(2)
  if (h==0)
legend("bottomleft", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
  box()
}

#-------------------
# Compute p-values of direct and HP-c forecasts: rely on HAC adjusted Wald-statistic
# -There is no HAC-adjustment for F-statistic
# -Use Wald-test with HAC-adjusted variance of estimators

ws_vec_unfiltered<-f_vec_unfiltered<-p_val_unfiltered<-MSE_unfiltered_vec<-NULL
for (h in 0:6)#h<-0
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected indicators on forward-shifted BIP: start at t=L (same sample as HP-C)
  lm_obj<-lm(forward_shifted_BIP[L:length(forward_shifted_BIP)]~x_mat_wc[L:nrow(x_mat_wc),])
# HAC adjustment
  var_HAC<-vcovHAC(lm_obj)
  coef<-lm_obj$coef
# Wald statistic  
  chisq<-t(coef)%*%solve(var_HAC)%*%coef
# P-Value  
  ws_vec_unfiltered<-c(ws_vec_unfiltered,pchisq(chisq,ncol(x_mat_wc)+1,lower.tail = F))
}

ws_vec_hp<-f_vec_hp<-t_mat_hp<-p_val_hp<-MSE_hp_vec<-NULL
for (h in 0:6)#h<-3
{
  # Shift BIP forward by publication lag+forecast horizon
  forward_shifted_BIP<-c(x_mat_wc[(1+lag_vec[1]+h):nrow(x_mat_wc),"BIP"],rep(NA,h+lag_vec[1]))
  # Regress selected filtered indicators on forward-shifted BIP
  length(forward_shifted_BIP)
  lm_obj<-lm(forward_shifted_BIP~t(hp_c_array_wc[,,1]))
# HAC adjustment  
  var_HAC<-vcovHAC(lm_obj)
  coef<-lm_obj$coef
# Wald statistic  
  chisq<-t(coef)%*%solve(var_HAC)%*%coef
# P-value  
  ws_vec_hp<-c(ws_vec_hp,pchisq(chisq,ncol(x_mat_wc)+1,lower.tail = F))

@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
cat("\\caption{Direct HP forecasts: entire data set with exclusion of the pandemic.", sep = "")
cat("\\label{direct_hp_forecasts}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
<<label=z_box_plot_pure_mba_2.pdf,echo=FALSE,results=tex>>=
file = "./Figures/direct_hp_forecasts_financial_crisis.pdf"
cat("\\begin{figure}[H]")
cat("\\begin{center}")
cat("\\includegraphics[height=5in, width=6in]{", file, "}\n",sep = "")
cat("\\caption{Shifted GDP (black), direct HP forecast (blue) and classic direct forecast (red) over the curse of the financial crisis:  regression based on all indicators and full sample information (without pandemic).", sep = "")
cat("\\label{direct_hp_forecasts_financial_crisis}}", sep = "")
cat("\\end{center}")
cat("\\end{figure}")
@
Moreover, the F-statistics are larger than for the direct forecasts, see Table \eqref{f_stat}, but the predictor remains statistically insignificant at forecast horizons exceeding two quarters (even though the evaluation is conducted in-sample and based on all available indicators).

@@

mat1<-rbind(round(ws_vec_hp,5),round(ws_vec_unfiltered,5))
colnames(mat1)<-paste("h=",0:(length(ws_vec_unfiltered)-1))
rownames(mat1)<-c("HP-C filtered","Unfiltered")
#latex(cor_vec, dec = 1, , caption = "Example of using latex to create table",
#center = "centering", file = "", floating = FALSE)
xtable(mat1, dec = 1,digits=rep(3,dim(mat1)[2]+1),
paste("Statistical significance of unfiltered and HP-C filtered forecasts based on HAC-adjusted Wald test: all indicators and full sample (without pandemic).  "),
label=paste("f_stat",sep=""),
center = "centering", file = "", floating = FALSE)
@

@@

par(mfrow=c(1,2))
colo<-c("black", "red", "green", "blue", "orange", "dimgray")
mplot<-gammak
plot(mplot[,1],main="M-MSE",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))

for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
axis(2)
legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
box()

mplot<-bk
plot(mplot[,1],main="M-SSA",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))

for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
axis(2)
box()
dev.off()

@@

par(mfrow=c(1,2))
colo<-c("black", "red", "green", "blue", "orange", "dimgray")
mplot<-bk
plot(mplot[,1],main="M-SSA: h=0",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))

for (i in 1:ncol(mplot))
{
  lines(mplot[,i],col=colo[i],lwd=1,lty=1)
}
abline(h=0)
axis(1,at=c(1,1+5*1:(nrow(mplot)/5)),labels=c(0,5*1:(nrow(mplot)/5)))
axis(2)
legend("topright", legend = colnames(mplot), col = colo, lty = 1, lwd = 1, bty = "n")
box()

bk<-matrix(mssa_indicator_obj$bk_x_array[select,select,,"h=4"],nrow=L)
colnames(bk)<-colnames(gammak)<-select_vec_multi
mplot<-bk
plot(mplot[,1],main="M-SSA: h=4",axes=F,type="l",xlab="",ylab="",col=colo[1],ylim=c(min(na.exclude(mplot)),max(na.exclude(mplot))))
